---
title: "p8105_hw2_as6447"
author: "Armaan Sodhi"
date: "2022-10-03"
output: github_document
---
##Problem 0
```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
```

```{r}
library(tidyverse)
library(readxl)
```

##Problem 1

Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Convert the entry variable from character (YES vs NO) to a logical variable (the ifelse or recode function may be useful).

```{r cars}
NYC_transit_data = 
  read_csv('data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv',
           col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>%
  janitor::clean_names()%>%
  select( 
          line, station_name, station_latitude, station_longitude, 
          entry, vending, exit_only, entrance_type, ada, starts_with('route'))%>%
  mutate(entry=if_else(entry=='YES',TRUE,FALSE))

nrow(NYC_transit_data)
ncol(NYC_transit_data)
```

From the use of `nrow` and `ncol` we find that there are 20 columns (variables) and 1868 rows (observations). It provides geographic data about where stations are located (ie station name, longitude/latitude), the characteristics of those locations (ie what type of entry do they have, are they ADA compliant, are there vending machines in the station), and the subway lines that run through the stations and the routes they follow. Cleaning the data required converting the variables to snake case (so that I can actually manipulate the variables without increasing my blood pressure), then I restricted the data set to only the required variables. Finally, I converted the entry variable to a logical variable. 

The data itself is not tidy considering the routes variables which are in `pivot_wider` so that each route is its own variable. It would be more efficient and "tidier" to `Pivot_longer`.


1.How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St 8th Avenue; 125st Broadway; 125st Lenox); the distinct function may be useful here.
```{r}
NYC_transit_data %>% distinct(line,station_name)
```
There are 465 unique stations (found in the console after running the above code chunk)

2.How many stations are ADA compliant?
```{r}
NYC_transit_data %>% filter(ada == TRUE) %>% distinct(line,station_name)
```

84 stations are found to be ADA Compliant. 

3.What proportion of station entrances / exits without vending allow entrance?
```{r}
NYC_transit_data %>% 
  filter(vending =='NO')%>%
  pull(entry)%>%
  mean()
```
The proportion of station entrances/exit without vending allow entrances is 0.377049 (37.70% of station entrances and exits are without vending allowed)

4.Reformat data so that route number and route name are distinct variables. 

How many distinct stations serve the A train? 

Of the stations that serve the A train, How many are ADA compliant? FINSIH

```{r}
NYC_transit_data %>%
  pivot_longer(
    NYC_transit_data, 
    route1:route11,
    names_to == 'route_name',
    values_to == 'route_number')

```

I reformatted the data so that the route number and names were distinct variables using the pivot_longer function. 

```{r}
NYC_transit_data %>% 
  filter(route_number == "A")%>%
  select(station_name,line) %>%
distinct

NYC_transit_data %>% 
  filter(route_number == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

There are 60 stations and lines that serve the A train and 17 stations that are ADA compliant.


##Problem 2

```{r pressure,}
Mr_Trash_Wheel = 
  readxl::read_excel("data/Trash Wheel Collection Data.xlsx", "Mr. Trash Wheel", range = "A2:N549")%>%
  janitor::clean_names()%>%
  drop_na()%>%
  mutate(
    sports_balls=round(sports_balls),
    sports_balls= as.integer(sports_balls),
    sheet_type= 'Mr_trash_wheel'
    )
```
This sheet initially has 548 observations and we have been given 16 variables, two of which where empty variables. We removed these variables
 and then ended up with only 14 variables. We omitted any rows that did not include dumpster-specific data. I then rounded and then converted it to an integer 
 
B.Use a similar process to import, clean, and organize the data for Professor Trash Wheel 
```{r}
Professor_Trash_Wheel = 
  readxl::read_excel("data/Trash Wheel Collection Data.xlsx", "Professor Trash Wheel", range = "A2:M96")%>%
  janitor::clean_names()%>%
  drop_na()%>%
  mutate(
    year = as.character(year),
    sheet_type= 'Professor_trash_wheel'
  )
```
Please note that I converted the year variable to a character here so that the two data sets are consistent with each other. I then created the `sheet_type` variable so that I could properly differentiate between the two when I merged them later. I dropped any rows that did not include dumpster-specific data as well. And conveted the variables to snake case.

```{r}
trash_wheel_full =
  full_join(Professor_Trash_Wheel, Mr_Trash_Wheel)
```

C. Write a paragraph about these data; You are encouraged to use inline R.
Be sure to note the number of observations in the resulting dataset
Give examples of key variables. 

In our new dataset there are 568 observations and 15 variables. Certain notable  

```{r}
trash_wheel_full %>% 
  filter(sheet_type == 'Professor_trash_wheel')%>%
  pull(weight_tons)%>%
  sum()
```
Utilizing `trash_wheel_bind %>% filter(sheet_type == 'Professor_trash_wheel')%>% pull(weight_tons)%>% sum()` it was found that the total weight of trash collected by Professor Trash Wheel was 162.54 tons. This was 
```{r}
trash_wheel_full %>% 
  filter(sheet_type == 'Mr_trash_wheel', year=='2020')%>%
  pull(sports_balls)%>%
```

In 2020 Mr. Trash Wheel collected 856 sports balls. 

To differentiate the data set 

##Problem 3

Use the data in pols-month.csv, unemployment.csv, and snp.csv. 

Our goal is to merge these into a single data frame using year and month as keys across datasets.

1.First, clean the data in pols-month.csv. 

Use separate() to break up the variable mon into integer variables year, month, and day; 

replace month number with month name; 

create a president variable taking values gop and dem, and 

remove prez_dem and prez_gop; and 

remove the day variable.
```{r}
pols_month_data = 
  read_csv('data/pols-month.csv')%>%
  janitor::clean_names()%>%
```

2.Second, clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.
```{r}
snp =
```
3.Third, tidy the unemployment data so that it can be merged with the previous datasets. This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.
```{r}
unemployment_data
```

4.Join the datasets by merging snp into pols, and merging unemployment into the result.
```{r}

```

5.Write a short paragraph about these datasets. Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).
