---
title: "p8105_hw2_as6447"
author: "Armaan Sodhi"
date: "2022-10-03"
output: github_document
---
##Problem 0
```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
```

```{r}
library(tidyverse)
library(readxl)
```

##Problem 1

Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Convert the entry variable from character (YES vs NO) to a logical variable (the ifelse or recode function may be useful).

```{r cars}
NYC_transit_data = 
  read_csv('data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv',
           col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>%
  janitor::clean_names()%>%
  select( 
          line, station_name, station_latitude, station_longitude, 
          entry, vending, exit_only, entrance_type, ada, starts_with('route'))%>%
  mutate(entry=if_else(entry=='YES',TRUE,FALSE))

nrow(NYC_transit_data)
ncol(NYC_transit_data)
```
Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and 

Give the dimension (rows x columns) of the resulting dataset. Are these data tidy?

SOLUTION: In the console I viewed the dataset and found what the dataset contains inside of it, with the total 


Answer the following questions using these data:
How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St 8th Avenue; 125st Broadway; 125st Lenox); the distinct function may be useful here.
How many stations are ADA compliant?
What proportion of station entrances / exits without vending allow entrance?
```{r}

```
SOLUTION:

Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?
```{r}

```

##Problem 2


```{r pressure,}
trash_wheel_data = read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx")
```

##Problem 3
