---
title: "p8105_hw2_as6447"
author: "Armaan Sodhi"
date: "2022-10-03"
output: git_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(readxl)
```

##Problem 1

```{r cars}
NYC_transit_data = read_csv('data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv')
```

Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Convert the entry variable from character (YES vs NO) to a logical variable (the ifelse or recode function may be useful).

Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?

Answer the following questions using these data:

How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St 8th Avenue; 125st Broadway; 125st Lenox); the distinct function may be useful here.
How many stations are ADA compliant?
What proportion of station entrances / exits without vending allow entrance?

Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?

##Problem 2

You can also embed plots, for example:

```{r pressure,}
trash_wheel_data = read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx")
```

##Problem 3
